### 2024-01-08 latex/figure5.tex

\begin{figure}[t]

\begin{minipage}[t]{\linewidth}

{\centering

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{from}\ClassTok{ simbio.io }\KeywordTok{import}\ClassTok{ biomodels, sbml}

\VariableTok{model }\OperatorTok{=}\ClassTok{ sbml}.\FunctionTok{load}\KeywordTok{(}\StringTok{"repressilator.sbml"}\KeywordTok{)}  \CommentTok{\# from existing SBML file}
\VariableTok{model }\OperatorTok{=}\ClassTok{ biomodels}.\FunctionTok{load\_model}\KeywordTok{(}\StringTok{"BIOMD12"}\KeywordTok{)}  \CommentTok{\# from BioModels}
\VariableTok{model}
\end{Highlighting}
\end{Shaded}

\begin{longtable}{@{}lll@{}}
    \multicolumn{3}{l}{Elowitz2000 - Repressilator} \\
    \toprule\noalign{}
    type & total & names \\
    \midrule\noalign{}
    \endhead
    \bottomrule\noalign{}
    \endlastfoot
    variables  &  6 & PX, PY, PZ, X, Y, Z \\
    parameters & 17 & cell, beta, alpha0, alpha, eff, n, KM, tau\_mRNA, tau\_prot, ... \\
    equations  & 12 & Reaction1, Reaction2, Reaction3, Reaction4, Reaction5, ... \\
\end{longtable}

}

\end{minipage}%

\caption{\label{fig-simbio-io}Creation of a model from a local SBML file
or one uploaded to BioModels.}

\end{figure}