### 2024-02-27 latex/figure6.tex
Use vector file for figure 6

\begin{figure}[t]

{\centering \includegraphics{src/performance/figures/performance}

}

\caption{\label{fig-runtime}Performance of different softwares to solve
models from the curated section of BioModels. (left) Run time for the
model BIOMD3 as a function of the number of output points. (right) Run
time for different models for 300 output points, using the geometric
mean of the different softwares to order them.}

\end{figure}