### 2024-03-21 article.tex
Rewrite introduction to clarify definition, translation and solving

\documentclass{article}
\usepackage[a4paper, margin=3cm, bottom=4cm]{geometry}

\usepackage{amsmath}
\usepackage[svgnames]{xcolor}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tcolorbox}

\usepackage{fancyvrb}
\newenvironment{CodeInput}
{\begin{tcolorbox}[title=input,boxrule=0pt]}
{\end{tcolorbox}}
\newenvironment{CodeOutput}
{\begin{tcolorbox}[title=output,boxrule=0pt]}
{\end{tcolorbox}}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.13,1.00}{#1}}
\newcommand{\ClassTok}[1]{\textcolor[rgb]{0.27,0.56,0.65}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.06,0.50}{#1}}
\newcommand{\ValueTok}[1]{\textcolor[rgb]{0.13,0.57,0.41}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.47,0.37,0.15}{#1}}
\newcommand{\IDETok}[1]{\textcolor[rgb]{0.58,0.58,0.58}{#1}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.06,0.50}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.70,0.27,0.27}{#1}}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph

\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}

\hypersetup{
  pdftitle={Poincaré and SimBio: a versatile and extensible Python ecosystem for modeling systems.},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
}

\usepackage{acronym}
\acrodef{CRN}[CRN]{chemical reaction network}
\acrodef{DSL}[DSL]{domain-specific language}
\acrodef{GPU}[GPU]{graphical processing unit}
\acrodef{GUI}[GUI]{graphical user interface}
\acrodef{IDE}[IDE]{integrated development environment}
\acrodef{JIT}[JIT]{just-in-time}
\acrodef{ODE}[ODE]{ordinary differential equation}
\acrodef{PyPI}[PyPI]{Python package index}
\acrodef{RHS}[RHS]{right-hand side}
\acrodef{SBML}[SBML]{System Biology Markup Language}
\acrodef{SDE}[SDE]{stochastic differential equation}
\acrodef{TPU}[TPU]{tensor processing unit}

\usepackage{authblk}
\title{Poincaré and SimBio: a versatile and extensible Python ecosystem for modeling systems.}
\author[1,2]{Mauro Silberberg}
\author[3]{Henning Hermjakob}
\author[3]{Rahuman S. Malik-Sheriff}
\author[1,2]{Hernán E. Grecco}
\affil[1]{Universidad de Buenos Aires, Facultad de Ciencias Exactas y Naturales, Departamento de Física. Buenos Aires, Argentina.}
\affil[2]{CONICET - Universidad de Buenos Aires, Instituto de Física de Buenos Aires (IFIBA). Buenos Aires, Argentina}
\affil[3]{European Bioinformatics Institute, European Molecular Biology Laboratory (EMBL-EBI), Wellcome Genome Campus, Cambridge, UK}

\begin{document}

\maketitle

\begin{abstract}
\Acf{CRN} play a pivotal role in diverse fields
such as systems biology, biochemistry, chemical engineering, and
epidemiology. High-level modelling of \acp{CRN} enables various simulation
approaches, including deterministic and stochastic methods. However,
existing Python tools for \ac{CRN} modelling typically wrap external C/C++
libraries for modelling and simulation, limiting their extensibility and
integration with the broader Python ecosystem. In response, we developed
Poincaré and SimBio, two novel Python packages for the definition and
simulation of dynamical systems and \acp{CRN}. Poincaré serves as a
foundation for dynamical system modelling, while SimBio extends this
functionality to \acp{CRN}, including support for the 
SBML. Poincaré and SimBio are developed as pure Python
packages enabling users to easily extend their simulation capabilities
by writing new or leveraging other Python packages. Moreover, this does
not compromise the performance, as code can be Just-In-Time compiled
with Numba. Our benchmark tests using curated models from the BioModels
repository demonstrate that these tools may provide a potentially
superior performance advantage compared to other existing tools.
Additionally, to ensure a user-friendly experience, our packages use
standard typed modern Python syntax that provides a seamless integration
with \acp{IDE}. Our Python-centric approach
significantly enhances code analysis, error detection, and refactoring
capabilities, positioning Poincaré and SimBio as valuable tools for the
modelling community.
\end{abstract}

\acresetall % reset acronyms after abstract

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\Acp{CRN} are a fundamental concept of modelling in numerous fields
including systems biology, biochemistry, chemical engineering and epidemiology.
They are comprised of a set of chemical species or biological entities
and a set of reactions that mediate transformations between them.
These systems can be simulated through multiple approaches:
deterministic \acp{ODE} to model macroscopic behavior,
\acp{SDE} to model microscopic fluctuations,
and jump processes (Gillespie-like simulations) to account for the discreteness of populations.
Instead of directly writing the equations for each of these formulations,
which is error-prone and difficult to reuse,
these models can be \emph{defined} in a higher-level description
that can be \emph{translated} into equations for the different types of simulations
and, then, \emph{solved} numerically.

Several tools already exist to \emph{define}, \emph{translate} and \emph{solve} \acp{CRN}.
BioSimulators.org \cite{shaikhBioSimulatorsCentralRegistry2022},
a registry of simulation tools,
lists at least 15 Python software including
COPASI \cite{hoopsCOPASICOmplexPAthway2006},
Tellurium \cite{choiTelluriumExtensiblePythonbased2018} and
PySB \cite{lopezProgrammingBiologicalModels2013}.
COPASI is a standalone software with a \ac{GUI}
that is widely used for its user-friendly interface and comprehensive features.
It also includes Python bindings \cite{bergmannBASICOSimplifiedPython2023b} that allow advanced scripting.
Tellurium is a Python-based modeling environment
that uses a C++ library called libRoadRunner in the backend to translate and solve models.
PySB is a Python library that created a \ac{DSL} using standard Python to define models,
which are then translated to \acp{ODE} using a Perl library called BioNetGen \cite{harrisBioNetGenAdvancesRulebased2016}.

One limitation of these tools is their extensibility from Python.
As they wrap libraries in other languages for defining, translating and/or solving models,
these steps cannot be altered or easily inspected from Python.
While they enable model definition and running simulations via Python scripts,
they cannot fully leverage Python's extensive package ecosystem.
For example,
COPASI and Tellurium do not allow the use of solvers defined in other Python packages,
and adding new integrators requires working with C++.
In particular, the step that translates into equations is not exposed by any of these tools.
As such, it is not possible to apply custom optimizations to the equations
or use automatic differentation packages such as JAX \cite{jax2018github} to compute the model's jacobian.

\input{latex/figure1.tex}

Another challenge is the way models are defined.
Many tools support the \ac{SBML} \cite{huckaSBMLL3V2} as an exchange format,
a \emph{de facto} standard for \acp{CRN} that defines species, parameters and reactions between species.
As writing \ac{SBML} directly is impractical,
Tellurium uses a \ac{DSL} called Antimony \cite{smithAntimonyModularModel2009} for defining models.
\Acp{DSL} allows to reuse the same code in different programming environments,
but are not recognized by default in \acp{IDE}
and, therefore, they cannot provide syntax highlighting, code completion, refactoring, and static analysis.
For Antimony, an extension providing this capabilities was developed for Visual Studio Code,
but its maintenance could be a demanding task for the systems biology community.
In the case of PySB,
using Python's dynamic nature,
it developers designed a \ac{DSL} within Python.
By default, it uses the global scope to create species and parameters,
without explicitly assigning them to Python variables or to the model,
but this approach is not fully compatible with \acp{IDE},
affecting the development experience.

To overcome these limitations,
we developed poincaré and SimBio,
open-source Python packages for defining, translating and solving systems.
Poincaré allows one to define differential equation systems
using variables, parameters and constants,
and assiging rate equations to variables.
For defining \acp{CRN},
SimBio builds on top of poincaré providing species and reactions that keep track of stoichiometries.
Both are focused on providing an ergonomic experience to end-users
by integrating well with \acp{IDE} and static analysis tools
through the use of standard modern Python syntax.
Moreover, since they are coded in pure Python,
each step from model definition, translation to equations or solving
can be extended or debugged from Python.
Being the first-ever pure Python packages for systems modelling,
they offer extensive extensibility,
from simple tasks like reusing integrators defined in other packages,
to complex ones like altering the compilation process to leverage some structure in the equations.
For example, using a for-loop in the compiled equations could improve the runtime performance
if there is some repetitive structure in the system,
as happens in spatial modelling.
The models built using these packages can be introspected to create other representations,
such as graphs connecting species and/or reactions, or tables with parameters or equations.
Furthermore, they have a modular architecture with a clear separation of concerns,
making it easier to maintain or to contribute new code,
which is beneficial for developers and maintainers.
We showcased the reliability of these tools by benchmarking them against the simulation results from other tools.
We also highlighted the substantial performance improvements our tools offer,
as this is crucial for construction and simulation of models of whole cells and organisms,
which necessitate the simulation of significantly large-scale models.

\hypertarget{results}{%
\section{Results}\label{results}}

\input{latex/figure2.tex}

Modular code architecture makes code reusable, extensible, and easier to
maintain. Therefore, we split the code to define and simulate reaction
systems into three Python packages: symbolite, to create symbolic
expressions; poincaré, to define dynamical systems; and simbio, to
define reaction systems and interface with systems biology standards
such as \ac{SBML}. These are pure Python packages with standard dependencies
from the PyData scientific stack such as NumPy \cite{harrisArrayProgrammingNumPy2020}
and pandas \cite{mckinneyDataStructuresStatistical2010}. They are published in the \acf{PyPI}
, where links to the source code and documentation hosted
in GitHub can be found, and can be easily installed with
\texttt{pip\ install\ \textless{}package\_name\textgreater{}}.

Symbolite is a lightweight symbolics package to create algebraic
mathematical expressions. Symbolite expressions can be inspected and
compiled to various backends. Currently, we have implementations for
NumPy \cite{harrisArrayProgrammingNumPy2020}; Numba \cite{lamNumbaLLVMbasedPython2015},
a \ac{JIT} compiler to LLVM; SymPy \cite{meurerSymPySymbolicComputing2017}, a
library for symbolic mathematics; and JAX \cite{jax2018github}, a
library that support automatic differentiation and compilation to \acp{GPU}
and \acp{TPU}. Symbolite is designed to facilitate the easy integration of
new backends.

\hypertarget{versatile-modelling-and-simulation-of-dynamical-systems-with-poincaruxe9}{%
\subsection{Versatile modelling and simulation of dynamical systems with
Poincaré}\label{versatile-modelling-and-simulation-of-dynamical-systems-with-poincaruxe9}}

\input{latex/figure3.tex}

Poincare is a package to define and simulate dynamical systems. It
provides a \texttt{System} class, where one can define
\texttt{Constant}s, \texttt{Parameter}s, \texttt{Variable}s, and create
equations linking a variable's derivative with an expression
(Figure~\ref{fig-first-order}). It also allows to define higher-order
systems by assigning an initial condition to a \texttt{Derivative}
(Figure~\ref{fig-second-order}).

Utilizing classes for system definition offers several advantages:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  The variable name to which a component is assigned can be
  automatically saved in the component for introspection (i.e.,
  \texttt{Oscillator.x.name\ ==\ "x"}),
\item
  It provides a namespace that allows to easily define multiple
  independent models in the same script,
\item
  It allows \acp{IDE} to provide autocomplete and refactoring capabilities
  (\texttt{Oscillator.\textless{}TAB\textgreater{}} shows \texttt{x},
  \texttt{v} and \texttt{eq}),
\item
  It allows creation of instances that can be composed into a bigger
  model (Figure~\ref{fig-composition}).
\end{enumerate}

\input{latex/figure4.tex}

For this last point, \acp{IDE} that support \texttt{dataclass\_transform} \cite{debontePEP681Data2021}
can provide a tooltip with the expected signature
(Figure~\ref{fig-ide}). This requires the use of type
annotations, which play a more significant role in static type checking
as they can help to identify errors before running the code. For
instance, to parameterize the initial conditions of variables we have to
use a \texttt{Constant}. If we try to use a \texttt{Parameter}, which
could be a time-dependent expression, it is flagged as a type error
(Figure~\ref{fig-ide}).

To simulate a system, we created a \texttt{Simulator} instance
(Figure~\ref{fig-sim}) that compiles the \ac{RHS} of a given system and interfaces
with \ac{ODE} solvers. By default, it creates a first-order
\ac{ODE} system using \texttt{numpy} as a backend. This can be
easily switched to other solvers. The Simulator wraps the output in a
\texttt{pandas.DataFrame}, which can be easily plotted with the standard
\texttt{plot} method.

Switching backends to \texttt{"numba"} can improve model runtime by a
factor up to x30 or more for big or long running models, however incurs
in a compile time penalty for the first run that must be taken into
account.

\hypertarget{extensible-definition-of-reaction-networks-using-simbio}{%
\subsection{Extensible definition of reaction networks using
SimBio}\label{extensible-definition-of-reaction-networks-using-simbio}}

For the reaction networks, our focus is on first-order differential
equations that describe the rate of change of species. SimBio simplifies
the definition of these network models by introducing \texttt{Species},
which consists of a \texttt{poincare.Variable} and a stoichiometric
number, and \texttt{RateLaw}, a construct that converts reactants into
products taking into account the stoichiometry
(Figure~\ref{fig-simbio}). Additionally, SimBio features
\texttt{MassAction}, a subclass of \texttt{RateLaw}, that intuitively
incorporates reactants into the rate law (Figure~\ref{fig-simbio}).

Several commonly used reactions are predefined as \texttt{MassAction}
subclasses, such as \texttt{MichaelisMenten}
(\(S + E \leftrightarrow ES \rightarrow P + E\)) and its approximate
form without the intermediate species \(ES\), and it is also simple to
implement your own as subclasses of \texttt{RateLaw} or
\texttt{MassAction}. Additionally, SimBio supports importing from and
exporting to \ac{SBML}, and downloading them directly from BioModels
\cite{malik-sheriffBioModels15Years2020} (Figure~\ref{fig-simbio-io}).

\input{latex/figure5.tex}

\hypertarget{reproducibility-and-performance}{%
\subsection{Reproducibility and
performance}\label{reproducibility-and-performance}}

\input{latex/figure6.tex}

To evaluate SimBio's reproducibility, we analyzed curated \ac{SBML} models
from BioModels \cite{malik-sheriffBioModels15Years2020}. Among the first 100 curated
models from BioModels we selected 60 that did not contains events, as
SimBio doesn't support events yet. We simulated the selected models with
COPASI and used the simulated results as ground truth, and demonstrated
that SimBio reproduces the results with minor numerical differences
depending on the solver tolerance.

For performance testing, we ran simulations using COPASI,
Tellurium/RoadRunner, and SimBio. Within SimBio, both NumPy and Numba
backends were considered. The LSODA solver was used for COPASI and
SimBio, while for Tellurium the comparable CVODE solver was used. In all
cases, we used relative and absolute tolerances of \(10^{-6}\). We
measured three simulation stages: loading, the initial (cold) run, and
subsequent (warm) runs for each model.

For COPASI and Tellurium/RoadRunner, we noted that its runtime depended
on the number of evaluation points, something that does not seem to
happen with SimBio (Figure~\ref{fig-runtime}, left). While SimBio's
NumPy backend is slower than both COPASI and RoadRunner, we obtained an
order of magnitude speed-up using the numba backend putting it on par
with them. A user might have to consider the trade-off between
compilation and run times, as the compilation of the
\ac{RHS} code might take longer than the runtime itself, and not be worth
it for running the model only once. Another speed-up in the runtime can
be had by switching the LSODA \texttt{scipy} solver for a more efficient
\texttt{numbalsoda} implementation that avoids the Python interpreter
between each of the integration steps. This last combination beats all
other methods, which is also true for the other models we tested
(Figure~\ref{fig-runtime}, right).

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

In this article, we introduced a suite of Python packages we developed
for defining and simulating dynamical systems and reaction networks.
These packages are deeply integrated with
\acp{IDE}, enabling code analysis tools to identify errors
prior to execution and assist in refactoring and code completion. We
adopted standard modern Python syntax to ensure seamless IDE
integration, supported by the extensive Python community.

Our approach differs from previous tools in that both the model
definition and its compilation into an
\ac{ODE} function are entirely Python-based. This approach simplifies the
development of various simulation methods, including performance
enhancements that exploit specific model structures. Importantly, being
Python-based does not compromise performance compared to C/C++ tools, as
the \ac{ODE} functions can be \ac{JIT} compiled using Numba.

The inclusion of \ac{SBML} support facilitates the effortless reuse of models
created by the systems biology community, along with the vast collection
of public models hosted in the BioModels repository. The modular
architecture of these packages facilitates their reuse, enhancement, and
extension by the wider Python community. For instance, an individual
from outside the systems biology field could contribute a stochastic
integrator to poincaré, which would then be available in SimBio. This
clear separation of concerns also makes the packages more
comprehensible, lowering the barrier for contributing improvements or
new features. Such an architecture ensures their maintainability and
ongoing development well into the future.

\hypertarget{resources}{%
\section{Resources}\label{resources}}

Source code repositories for these packages can be found at
\url{https://github.com/maurosilber/poincare} and
\url{https://github.com/hgrecco/simbio}.

\bibliographystyle{plain}
\bibliography{bibliography.bib}

\end{document}