### 2024-04-03 latex/figure6.tex

\begin{figure}[t]

{\centering \includegraphics{src/performance/figures/performance}

}

\caption{
    \label{fig-runtime}
    Performance of different softwares to solve models from the curated section of BioModels.
    (left) Run time for the model BIOMD3 as a function of the number of output points.
    (right) Run time for different models for 300 output points,
    using the geometric mean of the different softwares to order them.
    Each point corresponds to the median of 20 runs,
    with a neglibible errorbar given by the interquartile range.
}

\end{figure}